---
title: "Summary Statistics and Descriptive Plots"
author: "Joël de Vries"
date: "`r format(Sys.Date(), '%d %B %Y')`"
output:
  pdf_document:
    toc: true
  html_document:
    theme: flatly
    toc: true
    toc_float: true
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo    = FALSE,   # do not show code in output
  message = FALSE,
  warning = FALSE,
  fig.width  = 12,
  fig.height = 8
)

library(dplyr)
library(readr)
library(ggplot2)
library(tidyr)
library(knitr)
```

```{r data-loading}
# Estimation sample: participants meeting both inclusion criteria
#   (1) >= 10 weeks of observations
#   (2) >= 500 videos per week on average
# Excluded: p05 (56.6/week), p10 (11.8/week), p19 (140.2/week)
panel_all <- read_csv("gen/analysis/weekly_panel_estimation.csv",
                      show_col_types = FALSE)

# Variables in long format for the plots
variable_labels <- c(
  discovery_ratio      = "Discovery Ratio\n(share of new creators)",
  gem_sessieduur_mins  = "Avg. Session Duration\n(minutes)",
  like_rate            = "Like-Rate\n(likes / videos watched)",
  zoek_intensiteit     = "Search Intensity\n(number of searches)"
)

panel_long <- panel_all %>%
  select(deelnemer, week,
         discovery_ratio, gem_sessieduur_mins,
         like_rate, zoek_intensiteit) %>%
  pivot_longer(
    cols      = c(discovery_ratio, gem_sessieduur_mins,
                  like_rate, zoek_intensiteit),
    names_to  = "variable",
    values_to = "value"
  ) %>%
  mutate(
    variable = factor(variable,
                      levels = names(variable_labels),
                      labels = variable_labels)
  )
```

# Data overview

The estimation sample contains **`r nrow(panel_all)` rows** (participant × week combinations)
for **`r n_distinct(panel_all$deelnemer)` participants**, after excluding three participants
who did not meet the inclusion criteria (minimum 10 weeks of observations and minimum
500 videos per week on average; excluded: p05, p10, p19).
The observation period runs from
**`r format(min(panel_all$week), "%d %B %Y")`** to
**`r format(max(panel_all$week), "%d %B %Y")`**.

```{r weeks-per-participant}
panel_all %>%
  count(deelnemer, name = "Number of weeks") %>%
  rename(Participant = deelnemer) %>%
  kable(caption = "Number of observation weeks per participant")
```

---

# Summary statistics

The four research variables summarized across all participants and weeks in the estimation sample.

```{r summary-stats}
panel_all %>%
  select(discovery_ratio, gem_sessieduur_mins, like_rate, zoek_intensiteit) %>%
  summarise(across(
    everything(),
    list(
      N      = ~sum(!is.na(.)),
      Mean   = ~round(mean(., na.rm = TRUE), 3),
      SD     = ~round(sd(.,   na.rm = TRUE), 3),
      Min    = ~round(min(.,  na.rm = TRUE), 3),
      Max    = ~round(max(.,  na.rm = TRUE), 3),
      `N missing` = ~sum(is.na(.))
    ),
    .names = "{.col}__{.fn}"
  )) %>%
  pivot_longer(everything(),
               names_to  = c("Variable", "Statistic"),
               names_sep = "__") %>%
  pivot_wider(names_from = Statistic, values_from = value) %>%
  mutate(Variable = recode(Variable,
    discovery_ratio     = "Discovery Ratio",
    gem_sessieduur_mins = "Avg. Session Duration (min.)",
    like_rate           = "Like-Rate",
    zoek_intensiteit    = "Search Intensity"
  )) %>%
  kable(caption = "Descriptive statistics per research variable")
```

---

# Descriptive plots

## Timeline per participant

Each line represents one participant. The timeline shows how the variables
develop over the observation period. Notable trends and outliers are
immediately visible.

```{r plot-timeline, fig.cap="Development of research variables per participant over time"}
ggplot(panel_long,
       aes(x = week, y = value,
           colour = deelnemer, group = deelnemer)) +
  geom_line(alpha = 0.7, linewidth = 0.6) +
  geom_point(size = 0.8, alpha = 0.5) +
  facet_wrap(~variable, scales = "free_y", ncol = 2) +
  scale_x_datetime(date_breaks = "3 months", date_labels = "%b %Y") +
  labs(
    title    = "Development of research variables per participant",
    subtitle = "Each line = one participant | Weekly values",
    x        = NULL,
    y        = "Value",
    colour   = "Participant"
  ) +
  theme_minimal(base_size = 11) +
  theme(
    strip.text       = element_text(face = "bold"),
    axis.text.x      = element_text(angle = 45, hjust = 1, size = 8),
    legend.position  = "bottom",
    legend.key.width = unit(1.5, "cm"),
    panel.grid.minor = element_blank()
  )
```

## Distribution per variable

Boxplots provide an overview of the distribution of each variable across all participants
and weeks. Points outside the whiskers (1.5 × IQR) are considered outliers.
Individual data points are shown semi-transparently to convey the density of the data.

```{r plot-distribution, fig.cap="Distribution of research variables (all participants × all weeks)"}
ggplot(panel_long,
       aes(x = variable, y = value)) +
  geom_boxplot(outlier.shape = NA,
               fill = "#D0E8F2", colour = "#2C7BB6",
               width = 0.5, linewidth = 0.6) +
  geom_jitter(width = 0.15, alpha = 0.25,
              size = 0.9, colour = "#2C7BB6") +
  facet_wrap(~variable, scales = "free", ncol = 2) +
  labs(
    title    = "Distribution of research variables",
    subtitle = "Estimation sample (17 participants) × all weeks | Points outside whiskers are outliers",
    x        = NULL,
    y        = "Value"
  ) +
  theme_minimal(base_size = 11) +
  theme(
    strip.text       = element_text(face = "bold"),
    axis.text.x      = element_blank(),
    axis.ticks.x     = element_blank(),
    panel.grid.minor = element_blank()
  )
```

---

# Outlier analysis

Participant-week combinations above the 99th percentile per variable.
These are values that warrant closer inspection to determine whether they are
substantively correct or indicative of data issues.

```{r outliers}
panel_all %>%
  select(deelnemer, week,
         discovery_ratio, gem_sessieduur_mins,
         like_rate, zoek_intensiteit) %>%
  pivot_longer(-c(deelnemer, week),
               names_to  = "variable",
               values_to = "value") %>%
  group_by(variable) %>%
  mutate(p99 = quantile(value, 0.99, na.rm = TRUE)) %>%
  filter(value > p99) %>%
  mutate(
    week     = format(week, "%Y-%m-%d"),
    value    = round(value, 3),
    p99      = round(p99, 3),
    variable = recode(variable,
      discovery_ratio     = "Discovery Ratio",
      gem_sessieduur_mins = "Avg. Session Duration (min.)",
      like_rate           = "Like-Rate",
      zoek_intensiteit    = "Search Intensity"
    )
  ) %>%
  arrange(variable, desc(value)) %>%
  select(
    Variable    = variable,
    Participant = deelnemer,
    Week        = week,
    Value       = value,
    `P99 threshold` = p99
  ) %>%
  ungroup() %>%
  kable(caption = "Outliers above the 99th percentile per variable")
```

## Notes per variable

**Avg. Session Duration** — Values up to ~93 minutes (p06) are substantively plausible.
Long sessions occur when a participant uses TikTok for an extended period. There is no
reason to remove these values.

**Like-Rate** — Values above 1.0 are mathematically impossible given the definition
(likes / videos watched), but can arise due to a timing asynchrony: a video is watched
in week A but only liked in week B (e.g. via a creator's profile page). Values are
capped at 1.0.

**Search Intensity** — p06 shows consistently high search intensity (up to 172 per week),
with an increasing trend towards the end of the observation period. This is
substantively plausible (active search behavior) but deviates strongly from the other
participants. When modeling, it may be worth considering a log-transformation of this
variable due to its strong right skew.
